\begin{thebibliography}{99}

\bibitem{Cadena2016}
C. Cadena, L. Carlone, H. Carrillo, Y. Latif, D. Scaramuzza, J. Neira, I. Reid, and J. J. Leonard, ``Past, Present, and Future of Simultaneous Localization and Mapping: Toward the Robust-Perception Age,'' {\it {IEEE} Trans. Robotics}, Vol. 32, No. 6, pp. 1309--1332, 2016.

\bibitem{Grisetti2010}
G. Grisetti, R. Kümmerle, C. Stachniss, and W. Burgard, ``A Tutorial on Graph-Based {SLAM},'' {\it IEEE Intell. Transport. Syst. Mag.}, Vol. 2, No. 4, pp. 31--43, 2010.

\bibitem{Kuemmerle2011}
R. Kümmerle, G. Grisetti, H. Strasdat, K. Konolige, and W. Burgard, ``{g2o}: A General Framework for Graph Optimization,'' {\it Proc. {IEEE} Int. Conf. Robot. Autom. ({ICRA})}, pp. 3607--3613, 2011.

\bibitem{Kaess2012}
M. Kaess, H. Johannsson, R. Roberts, V. Ila, J. J. Leonard, and F. Dellaert, ``{iSAM2}: Incremental Smoothing and Mapping Using the Bayes Tree,'' {\it Int. J. Robotics Research}, Vol. 31, No. 2, pp. 216--235, 2012.

\bibitem{Cunningham2013}
A. Cunningham, V. Indelman, and F. Dellaert, ``{DDF-SAM} 2.0: Consistent Distributed Smoothing and Mapping,'' {\it Proc. {IEEE} Int. Conf. Robot. Autom. ({ICRA})}, pp. 5220--5227, 2013.

\bibitem{Choudhary2015}
S. Choudhary, L. Carlone, H. I. Christensen, and F. Dellaert, ``Exactly Sparse Memory Efficient {SLAM} using the Multi-Block Alternating Direction Method of Multipliers,'' {\it Proc. {IEEE/RSJ} Int. Conf. Intelligent Robots and Systems ({IROS})}, pp. 1349--1356, 2015.

\bibitem{Boyd2011}
S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein, ``Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers,'' {\it Found. Trends Mach. Learn.}, Vol. 3, No. 1, pp. 1--122, 2011.

\bibitem{Choudhary2017}
S. Choudhary, L. Carlone, J. Nieto, J. Rogers, H. I. Christensen, and F. Dellaert, ``Distributed Mapping with Privacy and Communication Constraints: Lightweight Algorithms and Object-Based Models,'' {\it Int. J. Robotics Research}, Vol. 36, No. 12, pp. 1286--1311, 2017.

\bibitem{Lajoie2020}
P.-Y. Lajoie, Y. Hu, Y. Chang, B. Ramtoula, and L. Carlone, ``{DOOR-SLAM}: Distributed, Online, and Outlier Resilient {SLAM} for Robotic Teams,'' {\it {IEEE} Robotics Autom. Lett.}, Vol. 5, No. 2, pp. 1656--1663, 2020.


\bibitem{Montemerlo2002}
M. Montemerlo, S. Thrun, D. Koller, and B. Wegbreit, ``{FastSLAM}: A Factored Solution to the Simultaneous Localization and Mapping Problem,'' {\it Proc. 18th Nat. Conf. on Artificial Intelligence ({AAAI})}, pp. 593--598, 2002.

\bibitem{Liu2016}
Q. Liu and D. Wang, ``Stein Variational Gradient Descent: A General Purpose Bayesian Inference Algorithm,'' {\it Proc. 30th Conf. Neural Information Processing Systems ({NIPS})}, pp. 2370--2378, 2016.

\bibitem{Pavlasek2023}
J. Pavlasek, J. J. Z. Mah, R. Xu, O. C. Jenkins, and F. Ramos, ``Stein Variational Belief Propagation for Multi-Robot Coordination,'' {\it arXiv:2311.16916}, 2023.

\bibitem{Koide2024MegaParticles}
K. Koide, S. Oishi, M. Yokozuka, and A. Banno, ``MegaParticles: Range-based 6-DoF Monte Carlo Localization with GPU-Accelerated Stein Particle Filter,'' {\it Proc. {IEEE} Int. Conf. Robot. Autom. ({ICRA})}, 2024.

\bibitem{Cao2024}
H. Cao, S. Shreedharan, and N. Atanasov, ``Multi-Robot Object {SLAM} Using Distributed Variational Inference,'' {\it arXiv:2404.18331}, 2024.

\bibitem{Agarwal2013}
P. Agarwal, G. D. Tipaldi, L. Spinello, C. Stachniss, and W. Burgard, ``Robust Map Optimization Using Dynamic Covariance Scaling,'' {\it Proc. {IEEE} Int. Conf. Robot. Autom. ({ICRA})}, pp. 62--69, 2013.

\bibitem{Sunderhauf2012}
N. Sünderhauf and P. Protzel, ``Switchable Constraints for Robust Pose Graph {SLAM},'' {\it Proc. {IEEE/RSJ} Int. Conf. Intelligent Robots and Systems ({IROS})}, pp. 1879--1884, 2012.

\bibitem{Olson2013}
E. Olson and P. Agarwal, ``Inference on Networks of Mixtures for Robust Robot Mapping,'' {\it Int. J. Robotics Research}, Vol. 32, No. 7, pp. 826--840, 2013.

\bibitem{Yang2020}
H. Yang, P. Antonante, V. Tzoumas, and L. Carlone, ``Graduated Non-Convexity for Robust Spatial Perception: From Non-Minimal Solvers to Global Outlier Rejection,'' {\it {IEEE} Robotics Autom. Lett.}, Vol. 5, No. 2, pp. 1127--1134, 2020.

\bibitem{Latif2013}
Y. Latif, C. Cadena, and J. Neira, ``Robust Loop Closing Over Time for Pose Graph {SLAM},'' {\it Int. J. Robotics Research}, Vol. 32, No. 14, pp. 1611--1626, 2013.

\bibitem{Mangelson2018}
J. G. Mangelson, D. Dominic, R. M. Eustice, and R. Vasudevan, ``Pairwise Consistent Measurement Set Maximization for Robust Multi-Robot Map Merging,'' {\it Proc. {IEEE} Int. Conf. Robot. Autom. ({ICRA})}, pp. 2916--2923, 2018.

\bibitem{Hong2016}
M. Hong, Z.-Q. Luo, and M. Razaviyayn, ``Convergence Analysis of Alternating Direction Method of Multipliers for a Family of Nonconvex Problems,'' {\it SIAM Journal on Optimization}, Vol. 26, No. 1, pp. 337--364, 2016.

\bibitem{Wang2019}
Y. Wang, W. Yin, and J. Zeng, ``Global Convergence of ADMM in Nonconvex Nonsmooth Optimization,'' {\it Journal of Scientific Computing}, Vol. 78, No. 1, pp. 29--61, 2019.

\bibitem{Blanco2012ATO}
J.-L. Blanco, ``A tutorial on SE(3) transformation parameterizations and on-manifold optimization,'' {\it MAPIR Group Technical report \#012010}, University of Malaga, 2010 (Updated 2012).

\bibitem{Tian2021}
Y. Tian, K. Khosoussi, D. M. Rosen, and J. P. How, ``Distributed Certifiably Correct Pose-Graph Optimization,'' {\it {IEEE} Trans. Robotics}, Vol. 37, No. 6, pp. 2137--2156, 2021.

\end{thebibliography}
